---
title: "data_challenge"
author: "Don Johnson"
date: "6/26/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


## Intro

The devices report data from 5 measurements: rpm, voltage, current, motor temperature, and inlet temperature. 

Noise: There are many outliers due to noise. Some of these are bunched. 

Data is written at irregular intervals. The time difference between data points is a normal distribution with mean 10 and deviance 2. The shortest intervals are about 3.5 minutes and the longest intervals are about 16.5 minutes. 

Time to failure | hours | days 
---- | --- | --- 
min |   238.1 |   10
max | 26663.7 | 1111
avg |  8251.2 |  344


## Data Clean-up

Steps for Data Clean-up (found in noiseclean.R):
* Convert timestamp to POSIX time object
* Determine the normal range of the measurement using the 1st and 99th quantile
* Increase range with a 30% buffer on each side to capture rare spikes
* Replace outliers with a rolling median
* New variable: "TTF" Time to Failure in hours
* Remove Seasonal component to data. 
  - create sin and cos with period of 365 days
  - fit linear model to a measurement with sin/cos
  - take residuals of measurement (repeat for all 5)
* new Boolean variables: will it fail in 3 hours/days/weeks? 

(as a next step, express time to failure as an exponential such as, expTTF = exp(-TTF/72), and choose hours (e.g., 72) to obtain best fit)
  

```{r clean}
library(ggplot2)
library(dplyr)
library(lubridate)

setwd("/Users/dojo/data/tagup/data-science-challenge/data/train/")

source("./noiseclean.R")
myfiles <- grep("_rms", dir(), value = TRUE)
unitDF <- lapply(myfiles, clean_noise)

ttf <- lapply(unitDF, function(x) head(x$TTF, 1))
ttf <- unlist(ttf)
summary(ttf)

```

## Identify states via clustering 

I will cluster one unit's data and use those centers to create clusters for the rest of the data. I'm not sure if this is a reliable way to obtain consistent clustering

```{r}
measurements <- c("rpm", "motor_voltage", "motor_current", 
               "motor_temp", "inlet_temp")
clust1 <- kmeans(unitDF[[1]][,measurements],7, 
                 algorithm = "Lloyd", iter.max = 100)  # 7 states?
centers <- clust1$centers
ggplot(data=unitDF[[1]], aes(x=time, y=motor_voltage, 
         color=as.factor(clust1$cluster))) + geom_point()
```


Looks okay. Repeat clustering for all datasets. 

```{r, echo=FALSE, eval=FALSE}
totDF <- data.frame()
for(df in unitDF){
    df$cluster <- as.factor(kmeans(df[,measurements], 
                                  centers = centers, algorithm = "Lloyd", 
                                  iter.max = 100)$cluster)
    totDF <- rbind(totDF, data.frame(expTTF = df$expTTF, 
                                  state = df$cluster))
}
```

```{r}
unitDF <- lapply(unitDF, function(x) {
    x$cluster <- as.factor(kmeans(x[,measurements], 
                                  centers = centers, algorithm = "Lloyd", 
                                  iter.max = 100)$cluster)
    return(x)
    })

```


Next steps: 
* find 'states' associated with end of device lifetime
* tokenize state-chain to find state transitions associated with end of lifetime

